---
title: "lake_data_analysis"
format: html
editor: visual
---

## General questions in this study

1.  Are unicellular and multicellular algae ecologically differentiated?
2.  How are unicellular and multicellular algae coexist over spatial and temporal scales.
3.  do multicellular taxa sustain better than unicellular ones (or over larger environmental variance, temporal span and spatial span)

## Specific problems to address in this study using HMSC

1.  What environmental variables affect the presence of unicellular algae
2.  What environmental variables affect the presence of multicellular algae
3.  Are the environmental variables affect the presence of unicellular and multicellular algae different?
4.  What environmental variables affect the abundance of unicellular algae
5.  What environmental variables affect the abundance of multicellular algae
6.  Are the environmental variables affect the abundance of unicellular and multicellular algae different?
7.  Are environmental variables affecting the presence of unicellular and multicellular algae different from the environmental variables affecting the abundance of unicellular and multicellular algae.
8.  Are the presence of unicellular and multicellular algae differ temporally over the season
9.  Are the abundance of unicellular and multicellular algae differ temporally over the season
10. Are the presences of unicellular and multicellular algae differ spatially over the season
11. Are the abundances of unicellular and multicellular algae differ spatially over the season
12. Which traits show more differentiation? (Multicellularity, mucus layer or typical cell number)
13. Are biotic variables affecting the presence and abundance of unicellular and multicellular algae ?

## Result section

1.  start by describing taxa, morphology, in swedish lakes
2.  then describe he temporal and spatial distribution of uni and multi

## Summary

frame intro to hypothesis: multi wider niche?: ecological expansion vs. differentiation

model presence and absence first including important traits( multi, mucus, cell number)

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
